Locomotion and Whole-Body Motion Planning for Humanoid Robots: an Integrated Approach

### 2012-06-20 paper.tex
Add precision about the constrained rrt method
Add snapshots of the 'College de France' experiment. Minor corrections.

There are two ways to use motion planners to generate dynamically balanced robotic motions.
The more general one is to plan in a robot dynamic space. By taking into account both robot
configuration and velocity, motions that satisfy dynamic balance constraints can be generated
at a planning phase. When planning motion for humanoid robots, this is a particularly costly
approach, as the size of the space to explore is augmented with the robot velocity and footstep
positions. The other way is to first plan  a geometric path that can be approximated by a
dynamic trajectory in a second step. The approach we present in this paper falls into the second 
category. Some feasible dynamic motions are inherently impossible to compute with this kind
of approach. For example, jumping motions cannot be generated by a purely geometric planner.

@@

The equivalent random  extension on a constrained manifold $\mathcal{M}$, 
defined by the constraint function $f$,
starts from a valid configuration $q_{near} \in \mathcal{M}$, and extends the tree
towards a random configuration $q_{rand}$, while keeping the constraints
satisfied. Extension attempts orthogonal to $\mathcal{M}$ are useless,
as newly added edges have to be included  in $\mathcal{M}$. To extend in directions
that follow the directions of $\mathcal{M}$, we rely on the framework of 
jacobian-based prioritized inverse kinematics. Algorithm~\ref{alg:constrained} presents 
the adaptation of the classic extend function, and Fig.~\ref{fig:gikrrt} illustrates this
extension. The idea is to first project $q_{rand}$ on the tangent
space to $\mathcal{M}$  at $q_{near}$. Let us call the projected configuration $q_{rand}'$.
$q_{rand}'$ is obtained by one step of prioritized inverse kinematics optimization. Let 
$q_{rand}''$  be the result of a call to $\text{constraint-solver}(q_{rand}',f,\epsilon)$.  It is 
the projection
of $q_{rand}'$ on $\mathcal{M}$. Instead of extending the tree from $q_{near}$ towards 
$q_{rand}$, the algorithm tries to extend from $q_{near}$ towards 
$q_{rand}''$\footnote{The algorithm presented here gives a precise idea of the method, without 
  focusing on technical details. Readers interested in the actual implementation can
  refer to \url{https://github.com/laas/hpp-constrained} and 
  \url{https://github.com/laas/hpp-constrained-planner}, where the open-source 
  code corresponding to this algorithm is made available.}.

@@

Fig.~\ref{fig:shelf} shows the resulting walk motion and Fig.~\ref{fig:shelf-waist} 
shows the robot CoM trajectory during locomotion. Extension 3 presents a 
comprehensive video of this problem, including the motion execution on
the real robot HRP-2.

@@

\includegraphics[width=0.5\linewidth]{pics/shelves/waist-trajectory.png}

@@

>>>>

There are two main ways of using motion planners to generate dynamically balanced robotic motions.
The more general one is to plan in a robot dynamic space. By taking into account both robot
configuration and velocity, motions that satisfy dynamic balance constraints can be generated
at a planning phase. When planning motion for humanoid robots, this is a particularly costly
approach, as the size of the space to explore is augmented with the robot velocity and footstep
positions. The other way is to first plan  a geometric path that can be approximated by a
dynamic trajectory in a second step. The approach we present in this paper falls into the second 
category. Some feasible dynamic motions are inherently impossible to compute with this kind
of approach. For example, jumping motions cannot be generated by a purely geometric planner.

@@

The equivalent random  extension on a constrained manifold $\mathcal{M}$, 
defined by the constraint function $f$,
starts from a valid configuration $q_{near} \in \mathcal{M}$, and extends the tree
towards a random configuration $q_{rand}$, while keeping the constraints
satisfied. Extension attempts orthogonal to $\mathcal{M}$ are useless,
as newly added edges have to be included  in $\mathcal{M}$. To extend in directions
that follow the directions of $\mathcal{M}$, we rely on the framework of 
jacobian-based prioritized inverse kinematics. Algorithm~\ref{alg:constrained} presents 
the adaptation of the classic extend function, and Fig.~\ref{fig:gikrrt} illustrates this
extension. The idea is to first project $q_{rand}$ on the tangent
space to $\mathcal{M}$  at $q_{near}$. Let us call the projected configuration $q_{rand}'$.
$q_{rand}'$ is obtained by one step of prioritized inverse kinematics optimization. Let 
$q_{rand}''$  be the result of a call to $\text{constraint-solver}(q_{rand}',f,\epsilon)$.  It is 
the projection
of $q_{rand}'$ on $\mathcal{M}$. Instead of extending the tree from $q_{near}$ towards 
$q_{rand}$, the algorithm tries to extend from $q_{near}$ towards 
$q_{rand}''$\footnote{This presentation attempts to give a precise idea of the algorithm, without 
  focusing on technical details. Readers interested in the actual implementation can
  refer to \url{https://github.com/laas/hpp-constrained} and 
  \url{https://github.com/laas/hpp-constrained-planner}, where the corresponding open-source 
  code is available.}. While extending the tree,
the configurations along the new edge are automatically projected onto $\mathcal{M}$. These projections 
are not very costly if the edge is close to the constrained manifold.

@@

Fig.~\ref{fig:shelf} shows the resulting walk motion and Fig.~\ref{fig:shelf-waist} 
shows the robot CoM trajectory during locomotion. Extension 3 presents a 
comprehensive video of this problem, including the motion execution on
the real robot HRP-2 on stage. Fig.~\ref{fig:shelf-cdf} shows some snapshots 
taken from Extension 3.

@@

\includegraphics[width=0.4\linewidth]{pics/shelves/waist-trajectory.png}

@@

\includegraphics[width=0.19\linewidth]{pics/shelves-cdf/1.jpg}
\includegraphics[width=0.19\linewidth]{pics/shelves-cdf/2.jpg}
\includegraphics[width=0.19\linewidth]{pics/shelves-cdf/3.jpg}
\includegraphics[width=0.19\linewidth]{pics/shelves-cdf/4.jpg}
\includegraphics[width=0.19\linewidth]{pics/shelves-cdf/5.jpg}

\caption{Execution of the walking trajectory by HRP-2 on stage. The robot first goes
  to the shelves to release the ball, then comes back to a rest position.}

\label{fig:shelf-cdf}
\end{figure}
